---
title: "Networks_HW7"
author: "KiseokUchicago"
date: "2021-05-26"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=11, fig.height=9,
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```

## Homework7 (Local stability analysis)
## Coding assignment for ECEV 44500 Networks in Ecology and Evolution
Professor: **Mercedes Pascual**, **Sergio A. Alcala Corona**, **Shai Pilosof**, **Carlos Serván** \
Student: **Kiseok Lee**

```{r}
library(igraph)
library(tidyverse)
library(magrittr)
library(MASS)
library(RColorBrewer)

## theme for ggplot

mytheme <- theme_bw() + theme(
  legend.title  = element_text( size=17),
  #  legend.position = "bottom",
  # legend.direction = "horizontal",
  legend.key = element_blank(),
  legend.text  = element_text( size=17),
  panel.background = element_blank(),
  panel.grid = element_blank(),
  text = element_text( family="Helvetica", size=19),
  strip.text.x = element_text(family = "Helvetica", size = 10),
  strip.text.y = element_text(family = "Helvetica", size = 10),
  panel.border = element_rect( colour = "black", size=1.3),
  axis.ticks = element_line(size = 1.3),
  strip.background = element_rect( fill = "transparent", size = 1.3, colour = "black"  ),
  strip.text = element_text(size = 19)
)
```

## Introduction
In this lab we are going to model the community matrix M as a random matrix. We start with the work of May (1973), which addressed the simplest case. We then explore the results of Allesina and colleagues, probing at each step the influence of different ecological features on the local stability of the community.

## 1. May’s results

Generally, we can analyze the local stability of an equilibrium point of a system of S functions and S variables by linearizing the functions and obtaining the Jacobian matrix J evaluated at the equilibrium (notice that there can be more than one equilibrium point). This is represented by the community matrix M. May sampled the entries of M from a uniform distribution and assumed that the entries Mij are i.i.d random variables with mean μ=0 and variance σ=1. Each entry represents the effect of species j on the population of species i. Therefore, the diagonal elements Mii represent the effect of each species on itself (self-regulation). May set the diagonal to d=−1. We will see later that these assumptions can be relaxed.\

The eigenvalues of M characterize the dynamics of the system “close” to equilibrium. There are S eigenvalues, and these are distributed approximately uniformly within a circle with a radius √S. To obtain a radius equal to 1, we can divide the off-diagonal values of M by √S.\

```{r}
# Here sigma=1
S <- 500
M <- matrix(runif(S^2, min = -sqrt(3), max = sqrt(3)), S, S) # We use sqrt(3) so the sd of the uniform distribution will be 1)
diag(M) <- -1
e <- eigen(M, only.values = TRUE)$values ## compute the eigenvalues of M
data_eigenvalues <- tibble(x = Re(e), y = Im(e), S = S) ## data frame storage useful for plotting with ggplot

data_eigenvalues %>% ggplot(aes(x = x, y = y)) + geom_point() + mytheme + coord_fixed() + geom_vline(xintercept = -1, color='blue')
```

M will be locally stable when all the real parts of the eigenvalues are negative (max(Re(λ))<0. We can visualize the expected eigenvalues boundaries by drawing a circle whose perimeter determines these boundaries.

```{r}
expectation_girko <- tibble(x = rep(seq(-1,1,0.001), 2), 
                                type = rep(c(1, -1), each = length(x)/2), 
                                y = sqrt(1 - (x)^2) * type)
expectation_girko %<>% mutate(x = sqrt(S) * x,
                             y = sqrt(S) * y)                          
                            
data_eigenvalues %>% ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_path(data = expectation_girko, colour = "red") +
  mytheme +
  xlab(expression(Re(lambda))) +
  ylab(expression(Im(lambda)))+coord_fixed()+ geom_vline(xintercept = -1, color='blue')
```

The matrix in this section is not stable. How can you make it stable? Show.

## 2. Role of sigma and d
Until now we have only used May’s results. Now we can probe how playing with some features of M will change local stability. σ will change the radius of the circle and d will shift its center. Note that S−−√ normalizes the radius of the circle to be σ. The diagonal represents the self regulation of species. Intuitively, stronger self-regulation will stabilize the system (shift the center of the circle to the left). The determination of the rightmost eigenvalue will be a result of both d and σ.

```{r}
S <- 500
sigma_seq <- c(0.5, 1, 1.5) # A sequence of sigma values
diag_seq <- c(-3, 0, 3) # A sequence of values for the diagonal

# This function creates a random matrix with a given sigma and d, and returns the ev.
get_eigenvalues_sigma_d <- function(sigma, d){
  M <- matrix(runif(S^2, min = -sqrt(3)*sigma, max = sqrt(3)*sigma), S, S)/sqrt(S) # Normalize by square root of S
  M <- M + diag(S) * d # Shift the diagonal
  e <- eigen(M, only.values = TRUE)$values
  return(tibble(x = Re(e), y = Im(e), S = S, d = d, sigma = sigma))
}

# Get the ev of random matrices for each combination of sigma and d
data_ev_sigma_d <- NULL
for (sigma in sigma_seq){
  for (d in diag_seq){
    data_ev_sigma_d <- rbind(data_ev_sigma_d, get_eigenvalues_sigma_d(sigma, d))
  }
}
data_ev_sigma_d %>% arrange(desc(sigma)) %>% ggplot(aes(x = x, y = y, colour = factor(sigma))) +
  geom_point(size = 2.5, alpha = 0.5) +
  mytheme +
  xlab(expression(Re(lambda))) +
  ylab(expression(Im(lambda))) +
  scale_colour_discrete(name = expression(sigma)) +
  coord_fixed() +
  facet_grid(d~.) +
  scale_x_continuous(breaks = seq(from = -4.5, to = 4.5, by=1.5)) +
  geom_vline(xintercept = c(-1.5,1.5), linetype='dashed')
```

Given a matrix with S=250 and a self-regulation value of -1.5, what is the necessary variance of the interactions to make it stable? Show analytically and computationally.

## 3. Probability of stability
The stability bound (the condition for the rightmost eigenvalue to have negative real part) is given by √SC⋅σ<−d. Therefore, setting the diagonal to −√SC will set the threshold for stability at σ=1. So, for large S, the system is expected to be stable when σ<1.

```{r}
S_seq <- c(25, 50, 100) # A curve per network size
nsim <- 100 # Generate multiple random matrices for each S
sigma_seq <- seq(0.8, 1.2, 0.01) # vary sigma
C <- 0.15

# Function to create a matrix and test if its righmost ev is negative.
stable_matrix <- function(i){
  M <- matrix(runif(S^2, min = -sqrt(3)*sigma, max = sqrt(3)*sigma) * rbinom(S^2, size = 1, prob = C), S, S)
  diag(M) <- 0
  M <- M + diag(S) * d
  e <- eigen(M, only.values = TRUE)$values
  return (max(Re(e)) < 0) # The rightmost eigenvalue should be negative for the matrix to be locally-stable
}

# Calculate the probability of stability for each combination of network size and sigma.
prob_stability_df <- NULL
for (S in S_seq){
  print(S)
  d <- -sqrt(S * C) # Set the diagonal to sqrt(SC) to make sigma=1
  for (sigma in sigma_seq){
    num_stable <- 0
    for (i in 1:nsim){
      num_stable <- num_stable+stable_matrix() # If the matrix is stable at it ot the count.
    }
    p_stability <- num_stable/nsim
    prob_stability_df <- rbind(prob_stability_df, data.frame(S=S, sigma=sigma, p_stability=p_stability))
  }
}

prob_stability_df %>% ggplot(aes(x = sigma, y = p_stability, colour = factor(S), shape = factor(S))) +
  geom_point(size = 3) +
  geom_line() +
  mytheme +
  xlab(expression(sigma)) +
  ylab(expression(p[stab])) +
  scale_colour_discrete(name = "S") +
  scale_shape_discrete(name = "S") +
  geom_vline(xintercept = 1, linetype='dashed')
```

How does the probability of stability depend on the number of species? Why? Try to increase the number of species.

## 4. The role of mean

The mean “determines” if interactions are more likely to be mutualistic (positive mean) or antagonistic (negative mean). Adding a constant to M without perturbing the diagonal will move one eigenvalue to approximately μ(S−1)+d. Where this eigenvalue falls with respect to the original spectrum can influence the stability. The more positive the mean, the more positive the eigenvalue and the higher the probability to destabilize the system.

```{r}
S <- 500
mu_seq <- c(-0.1, 0.01, 0.1)

outliers <- tibble(x = (S - 1) * mu_seq-sqrt(S), y = 0, mu = mu_seq) # Estimation of where the ev will fall

data_eigenvalues_mean <- map(mu_seq, # map is instead of using for. it maps a function to each element of the vector mu_seq
                             function(mu){
                               M <- matrix(runif(S^2, min = -sqrt(3), max = sqrt(3)), S, S)
                               diag(M) <- -sqrt(S)
                               M <- M + mu - diag(S) * mu
                               e <- eigen(M, only.values = TRUE)$values
                               tibble(x = Re(e), y = Im(e), S = S, mu = mu)
                             }) %>% bind_rows()


data_eigenvalues_mean %>% ggplot(aes(x = x, y = y)) + 
  geom_point(size = 1.3, alpha = 0.5) + 
  mytheme + 
  xlab(expression(Re(lambda))) + 
  ylab(expression(Im(lambda))) + 
  facet_grid(mu ~.) +
  coord_fixed() +
  geom_point(data = outliers, size = 2, colour = "red")

```

Try to examine the role of the mean for a different distribution than uniform (in line M <- matrix(runif(S^2, min = -sqrt(3), max = sqrt(3)), S, S))

## 5. Role of correlation (ρ)
Until now, Mij was independent of Mji. But in ecology this is not necessarily the case. For example, assuming μ=0, negative corelation is like prey-predator (+/-) and positive is like competition (-/-) or mutualism (+/+). When ρ=0, the eigenvalues fall on a circle, which is what we have seen so far. When ρ≠0, the eigenvalues fall on an ellipse. Negative correlation (ρ<0) shrinks the real part, whereas positive correlation (ρ>0) shrinks the imaginary part. So a negative correlation tends to stabilize the system.

```{r}
S <- 300
rho_seq <- c(-0.75, 0, 0.75)

generate_eigenvalues_cor <- function(rho){
  M <- matrix(0, S, S)
  Sigma <- matrix(c(1, rho, rho, 1), nrow = 2) ## set 2x2 covariance matrix
  M <- diag(rnorm(S)) ## set the diagonal, notice thas as long as it has mean 0 it can have a variance
  for(i in seq(1, S - 1)) {
    for (j in seq(i + 1, S)){
      Z <- mvrnorm(1, c(0,0), Sigma) # Now sample from a bivariate normal distribution with vector of means (0,0) and covariance matrix Sigma
      M[i,j] <- Z[1]  ## Set the offdiagonal pairs of M
      M[j,i] <- Z[2]
    }
  }
  M <- M / sqrt(S) # Rescale EV
  e <- eigen(M, only.values = TRUE)$values
  data_frame(x = Re(e), y = Im(e), S = S, rho = rho)
}

data_eigenvalues_cor <- map(rho_seq, generate_eigenvalues_cor) %>% bind_rows()

data_eigenvalues_cor %>% ggplot(aes(x = x, y = y, colour = factor(rho))) +
  geom_point(size = 2.5, alpha = 0.5) +
  mytheme +
  xlab(expression(Re(lambda))) +
  ylab(expression(Im(lambda))) +
  scale_colour_discrete(name = expression(rho)) +
  coord_fixed() +
  scale_x_continuous(breaks = c(1 + rho_seq, -1 - rho_seq)) +
  scale_y_continuous(breaks = c(1 - rho_seq, -1 + rho_seq))
```

## 6. Role of structure
We have changed parameters that influence interaction weights, but M was not structured. That is, the interactions were distributed randomly. Let’s see how adding blocks (modules) affects local stability. In the following example, each module has a random structure. We have already seen that when μ≠0, one eigenvalue was shifted to μ(S−1)+d. Now, this quantity may have different values for each of the blocks because the size of the block is different, and because it is also affected by the sign of the modularity. In particular for two blocks, there would be at most two outliers given approximately by the eigenvalues of the matrix of means, which is a block structured matrix having in each block a constant value depicting the mean interaction strength between the respective groups.

```{r}
S <- 100
C <- 0.4
alpha <- 1/4
m <- S * alpha  ## size of the top module
sigma <- 1
mu_seq <- c(-1/4, 0, 1/4)
#n <- length(mu_seq)
rho_seq <- c(-1/4, -3/4, -0.5)
Q_seq <- c(-0.5, 0, 0.35) # This is the modularity

generate_K <- function(Q){
  K <- matrix(0, S, S) ## adjacency matrix
  Cw <- C * (1 + Q/(alpha^2 + (1-alpha)^2)) # within block connectance
  Cb <- C * (1 - Q/(2 * alpha * (1-alpha))) # between block connectance
  # Between-block interactions
  AB <- matrix(runif(m * (S-m)) < Cb, m , S - m) 
  # Within-block interactions
  K[1:m, 1:m] <- matrix(runif(m^2) < Cw, m, m)
  K[(m + 1):S, (m+1):S] <- matrix(runif((S - m)^2) < Cw, S - m, S - m)
  # Add the between-block interactions to K
  K[1:m, (m+1):S] <- AB
  K[(m+1):S, 1:m] <- t(AB)
  return(K)
}

generate_M <- function(mu, rho, K, Q){
  M <- matrix(0, S, S) ## Community matrix
  Sigma <- matrix(c(sigma^2, rho * sigma^2, rho * sigma^2, sigma^2), nrow = 2) # Covariance matrix for bivariate normal distribution
  # Fill in M with interaction values
  for(i in seq(1, S - 1)) {
    for (j in seq(i + 1, S)){
      if (K[i,j] > 0){
        Z <- mvrnorm(1, c(mu, mu), Sigma) #
        M[i,j] <- Z[1] 
        M[j,i] <- Z[2]
      }   
    }
  }
  return(M)
}
```
Let’s see how these matrices look like

```{r}
# Plot the matrices
K <- generate_K(0.35)
heatmap(K, Rowv = NA, Colv = NA, symm = F, scale = 'none', revC = T, col=brewer.pal(2,"RdBu"))
M <- generate_M(0, 0.5, K)
heatmap(M, Rowv = NA, Colv = NA, symm = F, scale = 'none', revC = T)
```

Now find the stability.
```{r}
data_modules <- NULL
data_outliers <- NULL
for (Q in Q_seq){
  K <- generate_K(Q)
  for (i in 1:length(mu_seq)){
    mu <- mu_seq[i]
    rho <- rho_seq[i]
    print (paste(Q, mu, rho))
    M <- generate_M(mu, rho, K, Q)
    # Get the eigenvalues of M
    e <- eigen(M, only.values = TRUE)$values
    data_modules <- rbind(data_modules, data_frame(x = Re(e), y = Im(e), S = S, mu = mu, rho = rho, Q = Q))

    ## the following calculates the eigenvalues for the matrix of means + a small correction
    ## given by the correlation
    Cw <- C * (1 + Q/(alpha^2 + (1-alpha)^2)) # within block connectance
    Cb <- C * (1 - Q/(2 * alpha * (1-alpha))) # between block conenctnve
    mu_w <- Cw * mu
    mu_b <- Cb * mu
    data_outliers <- rbind(data_outliers,
    tibble(x = S/2 * (mu_w + c(1,-1) * (sqrt((1 - 4*alpha * (1- alpha)) * mu_w^2 + 4 * alpha* (1-alpha) * mu_b^2))), correction = ifelse(x !=0, rho/x, 0), y = 0, S = S, mu = mu, rho = rho, Q = Q)) # This line is taken from the paper.
  }
}

data_modules %>% ggplot(aes(x = x, y = y, colour = factor(mu))) + 
    geom_point(size = 2.5, alpha = 0.5) +
    geom_point(data = data_outliers, aes(x = x + correction), colour = "red") + 
  mytheme + xlab(expression(Re(lambda))) + 
  ylab(expression(Im(lambda))) + 
  scale_colour_discrete(name = expression(mu))+
  coord_fixed() + 
  facet_grid(Q ~ mu, 
             labeller = label_bquote(cols = paste(mu," = ",.(mu), collapse = ""), 
                                     rows = paste(Q," = ",.(Q), collapse = "")))
```

What do you notice about the effect of modularity vs. anti-modularity on the outlier eigenvalues?

## 7. Practicing linearizing a system to evaluate the local stability of an equilibrium.

To get a sense for what is involved in obtaining the linear approximation of a nonlinear system close to equilibrium, consider the Lotka-Volterra competition system in the paper by Maynard et al. (Ecology Letters 2018). You need to first find the equilibrium vector for species abundances, say x∗. Then, you would compute the Jacobian matrix J and evaluate it at this equilibrium to obtain the community matrix.

## Last problem. 
A general expression for x∗ is given in the paper, but show the steps (only a couple of lines of simple algebra). Then show the analytical formula for the Jacobian J and the corresponding community matrix for the nxn system. \

Finally, explore the eigenvalues of the community matrix for different parameter sets computationally. For this, set the number of species n to some given value of your choice, choose the interaction parameters aij (i≠j) at random, and make all the growth rates ri=1 and the self-regulation parameters aii=−1. \
```{r}
n <- 500
A = matrix(runif(n^2),nrow=n, ncol=n )
diag(A) <- -1
r <- rep(1,n)
D_r <- diag(r)
# plot matrix A in heatmap
heatmap(A, Rowv = NA, Colv = NA, symm = F, scale = 'none', revC = T)

# 1. solution
x_star = solve(A, -r)
# 2. Community matrix
M = diag(x_star) %*% A /sqrt(n) # # Normalize by square root of n
# 3. Get eigenvalues
e <- eigen(M, only.values = TRUE)$values ## compute the eigenvalues of M
data_eigenvalues <- tibble(x = Re(e), y = Im(e), S = S) ## data frame storage useful for plotting with ggplot
# plot
data_eigenvalues %>% ggplot(aes(x = x, y = y)) + geom_point() + mytheme + coord_fixed() + geom_vline(xintercept = 0, color='blue')
```

For different samples of the interaction parameters, obtain the corresponding eigenvalues and plot them in the 2-D space of their real and imaginary parts. Where do they fall? Why?\

The values fall within the circle with radius of 0.5. It is circular because of the circular law. σ will change the radius of the circle and here σ is 0.5. Dividing the elements of community matrix M with √S normalizes the radius of the circle to be σ.



