---
title: "Networks_HW2"
author: "KiseokUchicago"
date: "2021-04-11"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=11, fig.height=9,
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```

## Homework2 (Basic structural metrics)
## Coding assignment for ECEV 44500 Networks in Ecology and Evolution
Professor: **Mercedes Pascual**, **Sergio A. Alcala Corona** \
Student: **Kiseok Lee** 

```{r}
# libraries
library(igraph)
library(bipartite)
library(dplyr)
library(ggplot2)
```

## 1. Introduction

There is an endless number of metrics for network structure in the general network science, and several more developed particularly for ecological networks. We will naturally focus on the most relevant and popular ones. It is very useful however to have a reference for the most popular ones. In the context of ecology, two papers are particularly useful: \

Dormann CF, Fründ J, Blüthgen N, Gruber B. Indices, graphs and null models: analyzing bipartite ecological networks.
The Open Ecology Journal. 2009;2: 7–24.\

Delmas E, Besson M, Brice M-H, Burkle LA, Dalla Riva GV, Fortin M-J, et al. Analysing ecological networks of species interactions: Analyzing ecological networks.
Biol Rev. 2019;94: 16–36.\

## 2. Unipartite networks metrics

```{r}
# load data
# First, load the data 
otago_nodes <- read.csv('data/Otago_Data_Nodes.csv')
otago_links <- read.csv('data/Otago_Data_Links.csv')
otago_web <- graph.data.frame(otago_links, vertices = otago_nodes, directed = T)

# food web
chesapeake_nodes <- read.csv('data/Chesapeake_bay_nodes.csv', header=F)
names(chesapeake_nodes) <- c('nodeId','species_name')

chesapeake_links <- read.csv('data/Chesapeake_bay_links.csv', header=F)
names(chesapeake_links) <- c('from','to','weight')

ches_web <- graph.data.frame(chesapeake_links, vertices = chesapeake_nodes, directed = T)

plot(ches_web, edge.width=log(E(ches_web)$weight)/2, layout=layout.circle)
```
Otago web: Mouritsen KN, Poulin R, McLaughlin JP, Thieltges DW. Food web including metazoan parasites for an intertidal ecosystem in New Zealand. Ecology. Wiley Online Library; 2011;92: 2006–2006., with a description here.\

Chesapeake bay: Baird D, Ulanowicz RE. The Seasonal Dynamics of The Chesapeake Bay Ecosystem. Ecol Monogr. Ecological Society of America; 1989;59: 329–364. Data set is from here

## 2.1 Degree and strength

### 2.1.1. Unweighted degree
```{r}
deg_dist_out <- igraph::degree(otago_web, mode = 'out')  # igraph calculates the out-degree distribution
deg_dist_in <- igraph::degree(otago_web, mode = 'in') # and also the in-degree distribution

# Now we use a dataframe to make the plot.
df <- data.frame(deg=c(deg_dist_out,deg_dist_in), direction=c(rep('out',length(deg_dist_in)),rep('in',length(deg_dist_in))))
ggplot(df, aes(deg, fill=direction)) + geom_histogram(alpha=0.3)
```

## Problem 1. Transform the Otago web to an adjacency matrix (based on the previous lab), and then calculate node in-degree and out-degree using the matrix (i.e. write a short code).

```{r}
# make adjacency matrix
A_ota <- as_adjacency_matrix(otago_web)
class(A_ota)
dim(A_ota)

# in-degree
in_degree <- rep(0,dim(A_ota)[1])
for (i in 1:dim(A_ota)[1]){
  in_degree[i] <- sum(A_ota[i,])
}
in_degree

# out-degree
out_degree <- rep(0,dim(A_ota)[2])
for (j in 1:dim(A_ota)[2]){
  out_degree[j] <- sum(A_ota[,j])
}
out_degree

# Now we use a dataframe to make the plot.
df_A <- data.frame(deg=c(out_degree,in_degree), direction=c(rep('out',length(out_degree)),rep('in',length(in_degree))))
ggplot(df_A, aes(deg, fill=direction)) + geom_histogram(alpha=0.3, binwidth=1)

```

### 2.1.2. Weighted strength
```{r}
# degree
deg_dist_out <- igraph::degree(ches_web, mode = 'out')
deg_dist_in <- igraph::degree(ches_web, mode = 'in')

df <- data.frame(deg=c(deg_dist_out,deg_dist_in), direction=c(rep('out',length(deg_dist_in)),rep('in',length(deg_dist_in))))
ggplot(df, aes(deg, fill=direction))+geom_histogram(alpha=0.3,binwidth=1)
```

Strength
```{r}
s_dist_out <- igraph::strength(ches_web, mode = 'out')
s_dist_in <- igraph::strength(ches_web, mode = 'in')

df <- data.frame(s=c(s_dist_out,s_dist_in),direction=c(rep('out',length(s_dist_in)),rep('in',length(s_dist_in))))
ggplot(df, aes(s, fill=direction))+geom_histogram(alpha=0.3)
```

## Problem 2. Can discrepancies in in-degree and out-degree (or strength) give us some ecological insight?
Yes. The out-degree indicates that the node's degree in higher if the node is consumed by a larger number of predators, whereas in-degree indicates that the node's degree is higher if the node in consuming a larger number of preys. Therefore, larger out-degree means that the node is a widely consumed prey and larger in-degree means that the node is more of a generalist predator. The discrepancy simply means that consuming behavior of predators show different tendency compared to resource's consumption range. \
However, for strength, there is lesser discrepancy. This means that when the interaction is weighted the discrepancy disappears.

## 2.2 Clustering coefficient
```{r}
# The ratio of the triangles connected to the vertex and the triples centered on the vertex.
transitivity(otago_web, type = 'local') 
histogram(transitivity(otago_web, type = 'local'))

# The ratio of the triangles and the connected triples in the whole graph
transitivity(otago_web, type = 'global') 

```

## Problem 3. What can a high local clustering coefficient indicate in a food web?
I could mean that the local interactions are denser and more interconnected.

## 2.3 Geodesics
Geodesics (shortest paths) are at the basis of many metrics, including some centralities. We can obtain a distance table matrix (between pairs of nodes).
```{r}
geodesics <- distances(otago_web, mode = 'all') # Assume an undirected graph.Because this is also an unweighted network, the matrix values indicate how many _hops_ it would take to arrive from one node to another in the shortest way.
#geodesics
geodesics[1:5,1:5]  # just visualize a sample of the Shortest-paths table of the whole network
```

## 2.4 Node centralities
There are many centrality metrics, each with dozens of variations. Let’s take a look at the few common ones. Note that it is common to normalize the centrality measures. This is particularly helpful when comparing across networks. Read more about that normalization in the help and in Delmas et al (2018).
```{r}
# Assume undirected networks
CC <- igraph::closeness(otago_web, mode = 'all', normalized = T) 
BC <- igraph::betweenness(otago_web, directed = F, normalized = T)
EC <- igraph::eigen_centrality(otago_web, directed = F, scale = T)
EC <- EC$vector

# Ploting the web, re-sizing nodes by centrality (e.g. the betweenness centrality). Try to do the same by using other centralities.
V(otago_web)$BC <- BC
par(mar=c(0,0,0,0))
plot(otago_web, vertex.size=BC*100, vertex.label=NA, edge.arrow.width=0.5, edge.arrow.size=0.5, edge.curved=0.5, layout=layout.circle)

# Use CC
V(otago_web)$CC <- CC
par(mar=c(0,0,0,0))
plot(otago_web, vertex.size=CC*10, vertex.label=NA, edge.arrow.width=0.5, edge.arrow.size=0.5, edge.curved=0.5, layout=layout.circle)

# Use EC
V(otago_web)$EC <- EC
plot(otago_web, vertex.size=EC*10, vertex.label=NA, edge.arrow.width=0.5, edge.arrow.size=0.5, edge.curved=0.5, layout=layout.circle)

```

## Problem 4. Try to calculate the centrality measures for directed food webs.

```{r}
# Directed
# Assume directed networks
CCd <- igraph::closeness(otago_web, mode = c('out'), normalized = T) 
BCd <- igraph::betweenness(otago_web, directed = T, normalized = T)
ECd <- igraph::eigen_centrality(otago_web, directed = T, scale = T)
ECd <- ECd$vector

# Use BCd
V(otago_web)$BCd <- BCd
par(mar=c(0,0,0,0))
plot(otago_web, vertex.size=BCd*100, vertex.label=NA, edge.arrow.width=0.5, edge.arrow.size=0.5, edge.curved=0.5, layout=layout.circle)

# Use CCd
V(otago_web)$CCd <- CCd
par(mar=c(0,0,0,0))
plot(otago_web, vertex.size=CCd*100, vertex.label=NA, edge.arrow.width=0.5, edge.arrow.size=0.5, edge.curved=0.5, layout=layout.circle)

# Use EC
V(otago_web)$ECd <- ECd
plot(otago_web, vertex.size=ECd*10, vertex.label=NA, edge.arrow.width=0.5, edge.arrow.size=0.5, edge.curved=0.5, layout=layout.circle)
```

## Problem 5. Can you think of an ecological interpretation for each of these measures? Some intuition is given in the references by Martín-González et al. (2010) and Delmas et al. (2019).
(All citing from Delmas et al. (2019) and Martín-González et al. (2010))
- Closeness centrality:  It is a global measure  in  that,  although  defined  at  the  species  level,  it accounts for the structure of the entire network. It is based on the shortest path length between pairs of species and thus indicates how rapidly/efficiently a node is likely to influence the overall network.\
- Betweenness centrality: This measure is ideal to study the influence of species loss  on  fragmentation  processes,  for  example nodes with high CB values are considered as module connectors in modular networks.\
- Eigenvector centrality: This centrality is akin to a simulation of flow across interactions, in which each species influences all of its partners simultaneously. It then measures the relative importance of species by assigning them a score on the basis that an interaction with more influential species contributes more to a species' score than the same interaction with a low‐scoring species (Allesina & Pascual, 2009).\
- Therefore, it could be said that each measures has it's own strength: closeness (reflects shortest path), betweenness (reflect fragmentation process/modular role), eigenvector (reflects how much the node is connected to influential species)\

## Problem 6. What do you think would be the correlation beween these measures? Test it!
```{r}
library("ggpubr")
# use undirected centralities
df_cen <- data.frame(BC=BC,CC=CC,EC=EC)

# histograms
histogram(df_cen$BC, breaks=20, main="Betweenness")
histogram(df_cen$CC, breaks=20, main="Closeness") # close to normal distribution
histogram(df_cen$EC, breaks=20, main="Eigenvector")

# Shapiro-Wilk normality test
shapiro.test(df_cen$BC) # W = 0.49555, p-value < 2.2e-16 (not normally distributed alpha 0.05)
shapiro.test(df_cen$CC) # W = 0.98342, p-value = 0.03119 (not normally distributed alpha 0.05)
shapiro.test(df_cen$EC) # W = 0.89281, p-value = 4.269e-10 (not normally distributed alpha 0.05)

# qqplots
ggqqplot(df_cen$BC, ylab = "Betweenness")
ggqqplot(df_cen$CC, ylab = "Closeness")
ggqqplot(df_cen$EC, ylab = "Eigenvector")

# since the centrality measures are not normally distributed, let's use  Kendall rank correlation coefficient
res1 <-cor.test(df_cen$BC, df_cen$CC,  method = "kendall")
res1
res2 <-cor.test(df_cen$BC, df_cen$EC,  method = "kendall")
res2
res3 <-cor.test(df_cen$CC, df_cen$EC,  method = "kendall")
res3

# BC and CC
ggscatter(df_cen, x = "BC", y = "CC", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "kendall",
          xlab = "Betweenness", ylab = "Closeness")
# BC and EC
ggscatter(df_cen, x = "BC", y = "EC", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "kendall",
          xlab = "Betweenness", ylab = "Eigenvector")
# CC and EC
ggscatter(df_cen, x = "CC", y = "EC", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "kendall",
          xlab = "Closeness", ylab = "Eigenvector")
```

CC and EC shows higher correlation than between BC-CC, BC-EC.

## Problem 7. Which are the most central species in the network? Do they have anything in common?
To know which nodes have the highest centrality, we will use all 3 measures.
According to the graph below, 5, 43, 100, 114, 2, 127, 102, 110 nodes are high candidates.\
The common thing is that they all have high Eigenvector centrality. This makes sense because Eigenvector centrality is contingent upon the importance (centrality) of each other.
```{r}
# install.packages("scatterplot3d")
library("scatterplot3d")
scatterplot3d(df_cen, pch = 16, color="steelblue", angle = 29, type="h")

s3d <- scatterplot3d(df_cen, pch = 16, color="steelblue", angle = 29, type="h")
text(s3d$xyz.convert(df_cen), labels = rownames(df_cen),
     cex= 0.8, col = "red")
```

## 2.4 n-node sub-graphs (aka motifs)
These are sub-structural patterns composing a network, typically only meaningful in directed networks. There are 16 kinds of 3-node subgraphs. See them by calling help: ?triad.census. Let’s plot them.
```{r}
par(mfrow=c(4,4), mar=c(.75,.75,.75,.75))
for (i in 0:15){ # note that counting starts at 0
  plot(graph_from_isomorphism_class(size = 3, number = i),
       edge.arrow.size = 0.4,
       edge.color='black',
       main = i + 1)
  box(col='red')
}
```
Each of these different classes of subgraphs can have an ecological meaning. For example, #3 means that two prey share a predator and #5 is a trophic chain. It is therefore insightful to quantify how many of each kind are in the food web.

```{r}
motifs(otago_web) #absolute numbers
motifs(otago_web)/count_motifs(otago_web) # Proportion
```

## Problem 8. What is the dominant sub-graph in the Otago food web? What is its ecological meaning?
```{r}
df_motif <- data.frame(relative_abun=motifs(otago_web)/count_motifs(otago_web))
df_motif <- tibble::rownames_to_column(df_motif, var='motif_n')
m1 <- ggplot(data=df_motif, aes(x=motif_n, y=relative_abun)) + geom_bar(stat="identity") + theme_minimal() + ggtitle("Otago Motif relative abundance \n")
m1
```
7 is the most abundant. The center node is consumed by 2 predators. This is exploitative competition. (kind of indirect effect, interaction chain)

## Problem 9. Calculate the proportion of sub-graphs for the Bahia Falsa food web. Are their relative frequencies in this food web different from those for the Otago food web?

```{r}
# Bahia Falsa food web
Bahia_Falsa_metadata <- openxlsx::read.xlsx('data/Dunne2013PBioSupData.xlsx',sheet=3)

# We'll use free living species only (the first two columns)
edge_bahia <- Bahia_Falsa_metadata[,c(1,2)] 
colnames(edge_bahia) <- c("Consumer","Resource")
dim(edge_bahia)
edge_bahia2 <- edge_bahia %>% filter(!is.na(Consumer)) %>% filter(!is.na(Resource))
head(edge_bahia2)
tail(edge_bahia2)
dim(edge_bahia2)

bahia_web <- graph.data.frame(edge_bahia2, vertices = NULL, directed = T)

motifs(bahia_web) #absolute numbers
motifs(bahia_web)/count_motifs(bahia_web) # Proportion

df_motif_b <- data.frame(relative_abun=motifs(bahia_web)/count_motifs(bahia_web))
df_motif_b <- tibble::rownames_to_column(df_motif_b, var='motif_n')
m2 <- ggplot(data=df_motif_b, aes(x=motif_n, y=relative_abun)) + geom_bar(stat="identity") + theme_minimal() + ggtitle("Bahia Falsa Motif relative abundance \n")
m2


```
Bahia Falsa distribution of motif's relative abundance are similar to Otago's food web.


## 3. Network projections
As described below, we can create a one-mode projection from the two-mode bipartite graph with a simple step from linear algebra, namely matrix multiplication. There are two possible one-mode projections, one onto the groups and the other onto the nodes, depending on the order of the matrix product.\
We can project the incidence matrix by multiplying it with its transpose. 

## 3.1. One way of projection
```{r}
# Original
networklevel(memmott1999, index='connectance')

# Project plants
# Recall that the logical step '> 0' converts the original matrix into a binary one. Then 'tcrossprod' performs thee multiplication of the matrix with its transpose, to get the number of shared pollinators  
plants_projected <- tcrossprod(memmott1999>0) 
# In addition, we set the diagonal elements equal to zero, because Pii is otherwise the number of groups that node i belongs to
diag(plants_projected) <- 0

g <- graph.adjacency(plants_projected, mode = 'undirected', weighted = T)
par(mar=c(0,0,0,0))
plot(g)
plot(g, vertex.size=6, vertex.label=NA, edge.color='black', edge.width=log(E(g)$weight), layout=layout.circle, main="Unipartite_plant")

qplot(E(g)$weight)
```

## Problem 10. Try to project the pollinators and plot the degree distribution of this projection (use your code for question 1).

```{r}
# Project pollinators
# Recall that the logical step '> 0' converts the original matrix into a binary one. Then 'tcrossprod' performs thee multiplication of the matrix with its transpose, to get the number of shared pollinators  
pollinators_projected <- tcrossprod(t(memmott1999)>0) 
# In addition, we set the diagonal elements equal to zero, because Pii is otherwise the number of groups that node i belongs to
diag(pollinators_projected) <- 0

g2 <- graph.adjacency(pollinators_projected, mode = 'undirected', weighted = T)
par(mar=c(0,0,0,0))
plot(g2)
plot(g2, vertex.size=6, vertex.label=NA, edge.color='black', edge.width=log(E(g)$weight), layout=layout.circle, main="Unipartite_pollinator")

qplot(E(g2)$weight)

# Degree distribution

# make adjacency matrix
A_pol <- as_adjacency_matrix(g2)
class(A_pol)
dim(A_pol)

# in-degree
in_degree <- rep(0,dim(A_pol)[1])
for (i in 1:dim(A_pol)[1]){
  in_degree[i] <- sum(A_pol[i,])
}
in_degree

# out-degree
out_degree <- rep(0,dim(A_pol)[2])
for (j in 1:dim(A_pol)[2]){
  out_degree[j] <- sum(A_pol[,j])
}
out_degree

# Now we use a dataframe to make the plot.
df_A <- data.frame(deg=c(out_degree,in_degree), direction=c(rep('out',length(out_degree)),rep('in',length(in_degree))))
ggplot(df_A, aes(deg, fill=direction)) + geom_histogram(alpha=0.3, binwidth=1)
```

## 3.2. Another projection method
bi-adjacency matrix of a bipartite graph. This is a symmetric (and somewhat redundant) matrix with 0s between species of the same trophic level.

```{r}
plants_projected <- as.one.mode(memmott1999, project = 'full')
dim(plants_projected)

# lower = g x g matrix for g plants / higher = n x n matrix for n pollinators
plants_projected <- as.one.mode(memmott1999, project = 'lower')
dim(plants_projected)
g3 <- graph.adjacency(plants_projected, mode = 'undirected', weighted = T)
plot(g3,layout=layout.circle, main='Unipartite_plant')
qplot(E(g3)$weight)
```

## Problem 11. What information do we lose when projecting?
We lose the information of which plant interacts with which pollinator (if we are projecting on to the plants) and vice versa for the projection to the pollinators.

## Problem 12. What can you say about the density of the projected networks?
```{r}
par(mfrow=c(1,2))

plot(g, vertex.size=6, vertex.label=NA, edge.color='black', edge.width=log(E(g)$weight), layout=layout.circle, main="Unipartite_plant")

plot(g2, vertex.size=6, vertex.label=NA, edge.color='black', edge.width=log(E(g)$weight), layout=layout.circle, main="Unipartite_pollinator")
```
By comparing these two graphs of unipartite projections of plants and pollinators, we can see that the density of the projected networs depends on the direction(partition, species) of projection. Pollinator's unipartite graph appears to be much more "dense" compared to that of plant's, because the links are concentrated on small portion of the pollinator species.

## Problem 13. What is the ecological interpretation of the projections of a host-parasite network?
If we project and create a network of the host, the node with the higher degree(or centrality) means that it could be the good vector for transmitting parasites. So, it could be called a bridge species of paracites. Higher centrality could also mean that the node could be interacting more and coming to contact more often with other hosts.

## 4. Bipartite networks

## 4.1 Some basic network metrics
Going further, Dormann et al. (2009, The Open Ecology Journal) classified bipartite network metrics into: \

Metrics based on unweighted links\ 
Metrics based on weighted links\ 

Here are a few examples (see comments in code to learn what these are)

```{r}
# Connectance
data(olesen2002flores) # Check out the help for information on this data set!
olesen2002flores_binary <- 1 * (olesen2002flores > 0) # Make the data binary (unweighted)
I <- nrow(olesen2002flores_binary) # Number of lower level species (e.g., hosts, plants)
J <- ncol(olesen2002flores_binary) # Number of higher level species (e.g., parasites, pollinators)
S <- I + J # Total number of species, aka: Network size
L <- sum(olesen2002flores_binary > 0) # Number of edges in the network
A_i <- rowSums(olesen2002flores_binary) # The degree of hosts
A_j <- colSums(olesen2002flores_binary) # The degree of parasites
C <- L / (I * J) # Connectance
```

Notice that the following clustering coeffcients are defined at the local level, that is, for each species.
```{r}
cc_high <- colSums(olesen2002flores_binary) / nrow(olesen2002flores_binary) # Clustering coefficient higher level (the number of realized links divided by the number of possible links for each species)
cc_low <- rowSums(olesen2002flores_binary) / ncol(olesen2002flores_binary) # Clustering coefficient lower level (the number of realized links divided by the number of possible links for each species)
```

In a weighted network we can use strength instead of degree:
```{r}
S_i <- rowSums(olesen2002flores) # Node strength of hosts
S_j <- colSums(olesen2002flores) # Node strength of parasites
```

## Problem 14. Try to calculate these metrics for the memmott1999 data.
```{r}
# Connectance
data(memmott1999) # Check out the help for information on this data set!
memmott1999_binary <- 1 * (memmott1999 > 0) # Make the data binary (unweighted)
I <- nrow(memmott1999_binary) # Number of lower level species (e.g., hosts, plants)
J <- ncol(memmott1999_binary) # Number of higher level species (e.g., parasites, pollinators)
S <- I + J # Total number of species, aka: Network size
L <- sum(memmott1999_binary > 0) # Number of edges in the network
A_i <- rowSums(memmott1999_binary) # The degree of hosts
A_j <- colSums(memmott1999_binary) # The degree of parasites
C <- L / (I * J) # Connectance
C

cc_high <- colSums(memmott1999_binary) / nrow(memmott1999_binary) # Clustering coefficient higher level (the number of realized links divided by the number of possible links for each species)
# cc_high
cc_low <- rowSums(memmott1999_binary) / ncol(memmott1999_binary) # Clustering coefficient lower level (the number of realized links divided by the number of possible links for each species)
# cc_low
S_i <- rowSums(memmott1999) # Node strength of hosts
S_j <- colSums(memmott1999) # Node strength of parasites
```

## Problem 15. Can you plot the degree and strength distributions?
```{r}
# Degree of hosts
histogram(A_i, breaks = 30, main = "Degree of hosts" )

# Degree of parasites
histogram(A_j, breaks = 30, main = "Degree of parasites" )

# Node strength of hosts
histogram(S_i, breaks = 30, main = "Strength of hosts" )

# Node strength of parasites
histogram(S_j, breaks = 30, main = "Strength of parasites" )
```

## 4.2 Built-in metrics in bipartite

```{r}
# Network level
nl_metrics <- networklevel(olesen2002flores)
nl_metrics

# group level -> problem 16

# Species level -> problem 17
```

## Problem 16. Try to calculate and access metrics at the species level. Use help to understand which metrics bipartite has implemented.
```{r}
sl_metrics <- specieslevel(olesen2002flores)
sl_metrics
# help(specieslevel)
```

## Problem 17. Try to calculate and access metrics at the group level. Use help to understand which metrics bipartite has implemented.
```{r}
gl_metrics <- grouplevel(olesen2002flores)
gl_metrics
# help(grouplevel)
```

## Problem 18. Choose a couple of interesting metrics we have not discussed and briefly explain what they are and what is their meaning in a particular context (ecology, social sciences etc).
(1) togetherness: Network integration refers to a process of building links between two networks so that they dissolve into a single unified network. Togetherness measures the proximity of these two networks as they integrate; this notion is fundamental to social networks as it is relevant to important concepts such as trust, coherence and solidarity.\
Source: Moskvina, A., & Liu, J. (2016, August). Togetherness: an algorithmic approach to network integration. In 2016 IEEE/ACM International Conference on Advances in Social Networks Analysis and Mining (ASONAM) (pp. 223-230). IEEE.

(2) discrepancy: the number of links one would have to move to achieve perfect nestedness
(Brualdi & Sanderson, 1999)

(3) robustness: the area under the extinction curve (Burgos et al., 2007)

(4) niche overlap: with options for distance metrics (Pielou, 1972; Hurlbert, 1978)

(5) generality: for the higher trophic level 
(6) vulnerability: for the lower level (Bersier et al., 2002)

(7) partner diversity: simply the mean Shannon diversity of interactions of each species in
a level;

(8) fd (or alternatively functional diversity), which is similar to niche overlap but computed as branch length of a cluster diagram of dissimilarity of resource use (Devoto et al.,2012).

(9) C score: C score, or ‘checkerboardness’, which averages the number of instances of 01/10-patterns
(i.e. exclusive occurrences) for all pairwise species combinations (Stone & Roberts, 1990)

(10) V ratio: the variance ratio of species numbers to interaction numbers within species of a
level (Schluter, 1984)

To close, one very cool implementation in bipartite allows one to examine the degree distributions. It tries to fit the distributions to examine if they are close to a power law. The function is ‘degreedistr’. We will leave this for later because it includes issues concerning fitting the cumulative distributions and comparing the fits. But if you are curious, go ahead and explore …

## 5. Suggested reading
Some papers which develop or explain specific metrics:

Bersier L-F, Banašek-Richter C, Cattin M-F. Quantitative Descriptors of Food-Web Matrices. Ecology. Ecological Society of America; 2002; 83:2394–2407.\
Blüthgen N, Menzel F, Blüthgen N. Measuring specialization in species interaction networks. BMC Ecol. 2006;6: 9.\
Martín González AM, Dalsgaard B, Olesen JM. Centrality measures and the importance of generalist species in pollination networks. Ecol Complex. 2010; 7:36–43.\
Poisot T, Canard E, Mouquet N, Hochberg ME. A comparative study of ecological specialization estimators. Methods Ecol Evol. 2012; 3:537–544.\
For more details on projections you can see Mark Newman’s book: Networks an introduction (Section 6.6).\














