---
title: "Networks_HW3"
author: "KiseokUchicago"
date: "2021-04-19"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=11, fig.height=9,
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```

## Homework3 (Community detection, modularity)
## Coding assignment for ECEV 44500 Networks in Ecology and Evolution
Professor: **Mercedes Pascual**, **Sergio A. Alcala Corona** \
Student: **Kiseok Lee** 

```{r}
# libraries
library(igraph)
library(bipartite)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
```

## 1. Community detection (a.k.a “modularity”)

## 1.1 Unipartite networks
community structure detection algorithms are different.

### 1.1.1. Undirected, unweighted
Let’s try to detect communities in the Chesapeake Food Web from the last lab. This is a quantitative food web but for simplicity at this first stage we will work with an undirected, unweighted graph.

```{r}
# Load the web
chesapeake_nodes <- read.csv('data/Chesapeake_bay_nodes.csv', header=F)
names(chesapeake_nodes) <- c('nodeId','species_name')
chesapeake_links <- read.csv('data/Chesapeake_bay_links.csv', header=F)
names(chesapeake_links) <- c('from','to','weight')
ches_web <- graph.data.frame(chesapeake_links, vertices = chesapeake_nodes, directed = T)

# Detect communities, note the as.undirected command
ches_web_unweighted <- ches_web
ches_web_unweighted <- as.undirected(ches_web_unweighted)
E(ches_web_unweighted)$weight <- 1

# implementing Louvain method
cl <- cluster_louvain(ches_web_unweighted) # Can also use the weights = NULL argument
class(cl) # the result is of class communities

# module membership & visualization
module_membership <- membership(cl)
cols <- data.frame(mem=unique(module_membership), col= brewer.pal(length(unique(module_membership)), 'Set1'))
V(ches_web_unweighted)$module_membership <- module_membership
V(ches_web_unweighted)$color <- cols$col[match(V(ches_web_unweighted)$module_membership, cols$mem)]
plot(ches_web_unweighted, vertex.color=V(ches_web_unweighted)$color, vertex.size=10, vertex.label=NA, edge.arrow.width=0.3, edge.arrow.curve=0.5)
```

## Problem 1. Can you identify the species in each module? Does the assignment make sense?
```{r}
plot(ches_web_unweighted, vertex.color=V(ches_web_unweighted)$color, vertex.size=10, edge.arrow.width=0.3, edge.arrow.curve=0.5)

v_red <- names(module_membership[module_membership==1])
v_blue <- names(module_membership[module_membership==2])
v_green <- names(module_membership[module_membership==3])

# red nodes
chesapeake_nodes %>% filter(nodeId %in% v_red)
# blue nodes
chesapeake_nodes %>% filter(nodeId %in% v_blue)
# green nodes
chesapeake_nodes %>% filter(nodeId %in% v_green)
```

Red nodes and Green nodes makes sense. Red node comprises of microorganism or aquatic animals that dwell in the waters. Green node seems to be bigger animals (top predators like fish). Blue node makes less sense with my prior knowledge. But module would consist of a food web that could include both predator and prey that are closely associated. Therefore, Green node could be artificial.

## Problem 2 & 3. Try a different community detection method, like cluster_edge_betweenness. Are there any differences in module assignments?
```{r}
# implementing cluster_edge_betweenness method
cl_eb <- cluster_edge_betweenness(ches_web_unweighted) # Can also use the weights = NULL argument
class(cl_eb) # the result is of class communities

# module membership & visualization
module_member <- membership(cl_eb)
cols <- data.frame(mem=unique(module_member), col= brewer.pal(length(unique(module_member)), 'Set1'))
V(ches_web_unweighted)$module_member <- module_member
V(ches_web_unweighted)$color2 <- cols$col[match(V(ches_web_unweighted)$module_member, cols$mem)]
plot(ches_web_unweighted, vertex.color=V(ches_web_unweighted)$color2, vertex.size=10, vertex.label=NA, edge.arrow.width=0.3, edge.arrow.curve=0.5)
```
Yes, they are very different from the previouse Louvain's method. There are up to 7 modules compared to 3 modules before. Also, module 1 is dominant in number of nodes includes and other modules are relatively small, some modules even containing only 1 nodes.

### 1.1.2. Undirected, weighted

```{r}
cl_wt <- cluster_louvain(as.undirected(ches_web), weights = E(ches_web)$weight) #Notice the `weights` argument.
class(cl_wt) # the result is of class communities

module_membership_wt <- membership(cl_wt)
cols <- data.frame(mem=unique(module_membership_wt), col= brewer.pal(length(unique(module_membership_wt)), 'Set1'))
V(ches_web)$module_membership_wt <- module_membership_wt
V(ches_web)$color_wt <- cols$col[match(V(ches_web)$module_membership_wt, cols$mem)]
plot(as.undirected(ches_web), vertex.color=V(ches_web)$color_wt, vertex.size=5, vertex.label=NA, edge.arrow.width=0.3, edge.arrow.curve=0.5)

```

## Problem 4. Are the module assignments the same as in the unweighted network?
```{r}
par(mfrow=c(1,2))
plot(ches_web_unweighted, vertex.color=V(ches_web_unweighted)$color, vertex.size=10, vertex.label=NA, edge.arrow.width=0.3, edge.arrow.curve=0.5, main="Unweighted")

plot(as.undirected(ches_web), vertex.color=V(ches_web)$color_wt, vertex.size=5, vertex.label=NA, edge.arrow.width=0.3, edge.arrow.curve=0.5, main="Weighted")
```
No, they are not the same. The number of modules is also different.

Directed networks
We will not touch upon directed networks in the context of modularity maximization, although there is a development for it by Guimera et al (2007). The description of the sofware and the associated papers is here. The problem is that it is not available in an easy package, so you will have to hack your way through this. In fact, for directed networks we recommend using Infomap.

## 3. Bipartite networks
The difference between bipartite and unipartite networks lies in the “null” term of the modularity function, which should account for the fact that edges cannot be expected between nodes of the same set. This is well explained in (Guimera et al. 2007; Barber 2007) and reviewed in the context of ecology by Thebault (2013). Each of these have their own software associated with their papers. \

But for the sake of this exercise we are looking for an implementation to optimize modularity maximization which will be readily available, and also will allow us to analyze batches of networks (for testing patterns against shuffled networks). The bipartite package has a recent implementation of Stephen Beckett’s DIRTLPAwb+ algorithm (Beckett 2016). This algorithm maximizes a modularity function for weighted bipartite networks (Dorman & Strauss 2014), which is a recent extention of Barber’s (2007) method and is reduced to it if the input is a binary network. All this can be a bit confusing so it is strongly suggested to read these papers if you are about to apply these methods for community detection in your research.

```{r}
mod <- computeModules(memmott1999)
slotNames(mod) # see ?moduleWeb for details

mod@likelihood # This is the value of the modularity function Q. NOTICE THE @ SIGN (instead of $).
# 0.3042993

module_list <- listModuleInformation(mod) # The output is rather cumbersome...
par(mfrow=c(1,1))
plotModuleWeb(mod)

module_list <- module_list[[2]] # let's look at the modules. The first element in the list is the whole network, so start with 2
for (i in 1:2){ # Show the two first modules.
  message(paste('Module:',i))
  print(module_list[[i]])
}

```
This list data structure is horrible for analysis. This code transforms it to a data frame:
```{r}
# Transform the lists to data frames
m <- length(module_list) # Number of modules
mod_plants <- unlist(module_list, recursive = F)[seq(1,2*m,2)] # Assignments of plants
names(mod_plants) <- 1:m
mod_pollinators <- unlist(module_list, recursive = F)[seq(2,2*m,2)] # Assignments of pollinators
names(mod_pollinators) <- 1:m
tmp_plants <- data.frame(module = rep(names(mod_plants), sapply(mod_plants, length)), species = unlist(mod_plants), type='Plants')
tmp_poillnators <- data.frame(module = rep(names(mod_pollinators), sapply(mod_pollinators, length)), species = unlist(mod_pollinators), type='Pollinators')
# Make one data frame
(module_assignments <- rbind(tmp_plants,tmp_poillnators))

```

## Problem 5. Try to calculate modularity for the binary version of this web. How do the results differ?

```{r}
class(memmott1999)
dim(memmott1999)

# conver to binary matrix
sum(memmott1999!=0)
mem_binary <- ifelse(memmott1999>0, 1, 0)
sum(mem_binary!=0) # conversion worked well

mod_b <- computeModules(mem_binary)
slotNames(mod_b) # see ?moduleWeb for details

mod_b@likelihood # This is the value of the modularity function Q. NOTICE THE @ SIGN (instead of $).
# 0.3284303

module_list <- listModuleInformation(mod_b) # The output is rather cumbersome...
par(mfrow=c(1,1))
plotModuleWeb(mod_b)

module_list <- module_list[[2]] # let's look at the modules. The first element in the list is the whole network, so start with 2
for (i in 1:2){ # Show the two first modules.
  message(paste('Module:',i))
  print(module_list[[i]])
}
# This list data structure is horrible for analysis. This code transforms it to a data frame:

# Transform the lists to data frames
m <- length(module_list) # Number of modules
mod_plants <- unlist(module_list, recursive = F)[seq(1,2*m,2)] # Assignments of plants
names(mod_plants) <- 1:m
mod_pollinators <- unlist(module_list, recursive = F)[seq(2,2*m,2)] # Assignments of pollinators
names(mod_pollinators) <- 1:m
tmp_plants <- data.frame(module = rep(names(mod_plants), sapply(mod_plants, length)), species = unlist(mod_plants), type='Plants')
tmp_poillnators <- data.frame(module = rep(names(mod_pollinators), sapply(mod_pollinators, length)), species = unlist(mod_pollinators), type='Pollinators')
# Make one data frame
(module_assignments2 <- rbind(tmp_plants,tmp_poillnators))
```

```{r}
par(mfrow=c(1,2))
plotModuleWeb(mod)
plotModuleWeb(mod_b)
par(mfrow=c(1,1))

# module_assignments
# module_assignments2

module_assignments %>% filter(type=="Plants")
module_assignments2 %>% filter(type=="Plants")

```

(1) The modularity Q increased about 10% with binary matrix: from 0.3042993 to 0.3284303
(2) The assignment of the nodes to the module is also slightly different. For example, for plants the number of modules increased by 1 (5 -> 6) in binary matrix. In addition, the biggest module's size increased to 11 from 9.  

## Problem 6. Can you project the network and calculate modularity in each of the projections? Do plants that cluster together in the projection also cluster together in the bipartite version? Is there an difference in the ecological interpretation between these two scenarios?

```{r}
# lower = g x g matrix for g plants / higher = n x n matrix for n pollinators
plants_projected <- as.one.mode(memmott1999, project = 'lower')
dim(plants_projected)
# plants_projected

# project to pollinators
# pollinator_projected <- as.one.mode(memmott1999, project = 'higher')
# dim(pollinator_projected)

rownames(plants_projected)
colnames(plants_projected)

isSymmetric.matrix(plants_projected) # symmetric, therefore undirected

g_plant <- igraph::graph.adjacency(plants_projected, mode = 'undirected')
par(mar=c(0,0,0,0))
plot(g_plant, vertex.size=6, edge.curved=0.5)

# Detect communities, note the as.undirected command
E(g_plant)$weight <- 1

# implementing Louvain method
cl <- cluster_louvain(g_plant) # Can also use the weights = NULL argument
class(cl) # the result is of class communities

# module membership & visualization
module_membership <- membership(cl)
cols <- data.frame(mem=unique(module_membership),col=c('blue','red'))

V(g_plant)$module_membership <- module_membership
V(g_plant)$color <- cols$col[match(V(g_plant)$module_membership, cols$mem)]
plot(g_plant, vertex.color=V(g_plant)$color, vertex.size=10, edge.arrow.width=0.3, edge.arrow.curve=0.5)

# compare membership
# (1) module membership for bipartite version
mod_bipartite <- module_assignments %>% filter(type=="Plants")
table(mod_bipartite$module)
# (2) module membership for unipartite version
table(module_membership)

# look for discrepencies
# (1) Module that Leontodon.autumnalis is in for bipartite case
Leon_mod <- mod_bipartite %>% filter(species=='Leontodon.autumnalis')
mem_bi <- mod_bipartite %>% filter(module==Leon_mod$module) %>% select(species) %>% unlist()

# (2) Module that Leontodon.autumnalis is in for unipartite modules
mod_unipartite <- data.frame(species=names(module_membership),module=as.numeric(module_membership))
Leon_mod_uni <- mod_unipartite %>% filter(species=='Leontodon.autumnalis') # Leontodon.autumnalis is in module 3
mem_uni <- mod_unipartite %>% filter(module==Leon_mod_uni$module) %>% select(species) %>% unlist() 

# see difference
setdiff(mem_bi,mem_uni)
setdiff(mem_uni,mem_bi)

```
(1) The number of modules is different between bipartite and unipartite modules
(2) For example for the module that Leontodon.autumnalis is in, bipartite module is smaller than that of unipartite version. They are not the exact subset of one another. So we can say that not all plants that cluster together in the projection also cluster together in the bipartite version. 
(3) There is a difference in the ecological interpretation between these two scenarios. We lose information when we project the adjacency matrix into unipartite networks. Therefore, when creating modules, we no longer take the pollinator information into account (directly), which is ecologically different than having the pollinator information to directly maximize modularity Q. This is why the unipartite network's modular structure becomes much simpler(less number of modules) than that of bipartite version.

## Problem 7. Find another data set that interests you personally in which you could hypothesize the existence of communities. Explain why this hypothesis is relevant for this data set. Analyze for communities, with a method from the material above. Do the results support the hypothesis (e.g., when you examine the nodes in each module).
I would be interested in the modularity of fungal-bacterial network of rice seed. Hypothesis is that fungi and bacteria would not exist as separate module in the network. microbiome.(https://microbiomejournal.biomedcentral.com/articles/10.1186/s40168-020-00805-0)
```{r}
# domesticated rice seed microbiome
domestic_links <- read.csv('data/2_0.3_node_dom.css.ss.5.tsv',sep="\t", header=T)
b_domestic_links <- read.csv('data/5_0.3_edge_dom.phy.ss.5.tsv',sep="\t", header=T)
f_domestic_links <- read.csv('data/8_0.3_edge_dom.fun.ss.5.tsv',sep="\t", header=T)

# wild rice seed microbiome
wild_links <- read.csv('data/3_0.3_edge_wil.css.ss.5.tsv',sep="\t", header=T)
b_wild_links <- read.csv('data/6_0.3_edge_wil.phy.ss.5.tsv',sep="\t", header=T) # bacteria
f_wild_links <- read.csv('data/9_0.3_edge_wil.fun.ss.5.tsv',sep="\t", header=T) # fungi

# microbial network of wild rice seed endophytes
wild_web <- graph.data.frame(wild_links, directed = F)
b_wild_web <- graph.data.frame(b_wild_links, directed = F) # bacteria
f_wild_web <- graph.data.frame(f_wild_links, directed = F) # fungi

plot(wild_web, vertex.size=5, vertex.label=NA, main="Bacteria & Fungi-Wild rice")
plot(b_wild_web, vertex.size=5, vertex.label=NA, main="Bacteria-Wild rice")
plot(f_wild_web, vertex.size=5, vertex.label=NA, main="Fungi-Wild rice")


# Detect communities, note the as.undirected command
wild_unweighted <- wild_web
wild_unweighted <- as.undirected(wild_unweighted)
E(wild_unweighted)$weight <- 1

# implementing Louvain method
cl <- cluster_louvain(wild_unweighted) # Can also use the weights = NULL argument
class(cl) # the result is of class communities

# module membership & visualization
module_membership <- membership(cl)
table(module_membership)

my_color_collection <- c(
  "#CBD588", "#5F7FC7", "orange", "#AD6F3B", "#673770", 
  "#D14285", "#652926", "#C84248", "#8569D5", "#5E738F",
  "#D1A33D", "#8A7C64", "#599861","#616163", "#FFCDB2",
  "#6D9F71", "#242F40",
   "#CCA43B", "#F92A82", "#ED7B84", "#7EB77F", 
  "#DEC4A1", "#E5D1D0", '#0E8482', '#C9DAEA', '#337357', 
  '#95C623', '#E55812', '#04471C', '#F2D7EE', '#D3BCC0', 
  '#A5668B', '#69306D', '#0E103D', '#1A535C', '#4ECDC4', 
  '#F7FFF7', '#FF6B6B', '#FFE66D', '#6699CC', '#FFF275', 
  '#FF8C42', '#FF3C38', '#A23E48', '#000000', '#CF5C36', 
  '#EEE5E9', '#7C7C7C', '#EFC88B', '#2E5266', '#6E8898', 
  '#9FB1BC', '#D3D0CB', '#E2C044', '#5BC0EB', '#FDE74C', 
  '#9BC53D', '#E55934', '#FA7921', "#CD9BCD", "#508578", "#DA5724")

# color
cols <- data.frame(mem=unique(module_membership), col= my_color_collection[1:length(unique(module_membership))])
V(wild_unweighted)$module_membership <- module_membership
V(wild_unweighted)$color <- cols$col[match(V(wild_unweighted)$module_membership, cols$mem)]

# shape: triangle fungi / circle bacteria
V(wild_unweighted)$shape <- ifelse(substr(V(wild_unweighted)$name,1,1)=="B","circle","csquare")

# plot network
plot(wild_unweighted, vertex.color=V(wild_unweighted)$color, vertex.size=6, edge.arrow.width=0.3, edge.arrow.curve=0.5, vertex.shape=V(wild_unweighted)$shape, vertex.label=NA)
```
Circle is bacteria and square is fungi. We can see that square and circle coexist in each module.




## 4. References
Barber M. Modularity and community detection in bipartite networks. Physical Review E. 2007;76: 1–9.\
Beckett SJ. Improved community detection in weighted bipartite networks. R Soc Open Sci. 2016;3: 140536.\
Dormann CF, Strauss R. A method for detecting modules in quantitative bipartite networks. Methods Ecol Evol. 2014;5: 90–98.\
Guimerà R, Sales-Pardo M, Amaral LAN. Module identification in bipartite and directed networks. Phys Rev E Stat Nonlin Soft Matter Phys. 2007;76: 036102.\
Thébault E. Identifying compartments in presence-absence matrices and bipartite networks: insights into modularity measures. J Biogeogr. 2013;40: 759–768.)\

To go deeper into the topic of Community Detection in networks, here are some useful references:\

Fortunato, Santo, and Darko Hric. “Community detection in networks: A user guide.” Physics reports 659 (2016): 1-44.\
Fortunato, Santo. “Community detection in graphs.” Physics reports 486, no. 3-5 (2010): 75-174.\


